### 2021-02-23 architecture.tex
Finalize the text for first submission

\end{tikzpicture}
\end{figure}

@@

\subsubsection{Cost minimization by sorting}
\begin{lemma}
Precedence cost $C$ is minimized by the precedence that sorts the symbols by their costs in non-increasing order:
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
where $\argsort^-(x)$ is the set of all permutations $\rho$ that sort vector $x$ in non-increasing order ($x_{\rho(1)} \geq x_{\rho(2)} \geq \ldots \geq x_{\rho(n)}$).
\end{lemma}

>>>>

\end{tikzpicture}
\end{figure}
\todo{FB: Clean up the diagram.}

@@

\subsubsection{Cost minimization by sorting}
\begin{lemma}
Precedence cost $C$ is minimized by the precedence that sorts the symbols by their costs in non-increasing order:
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
where $\argsort^-(x)$ is the set of permutations $\rho$ that sort vector $x$ in non-increasing order ($x_{\rho(1)} \geq x_{\rho(2)} \geq \ldots \geq x_{\rho(n)}$).
\end{lemma}

### 2021-02-23 introduction.tex
Finalize the text for first submission

Modern saturation-based Automatic Theorem Provers (ATPs) such as E \cite{SCV:CADE-2019}, SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09} 
or \Vampire{} \cite{DBLP:conf/cav/KovacsV13}
employ the superposition calculus \cite{DBLP:journals/logcom/BachmairG94,DBLP:books/el/RV01/NieuwenhuisR01} as their underlying inference system.
Integrating flavours of resolution \cite{DBLP:books/el/RV01/BachmairG01}, paramodulation \cite{Robinson1983}, and 
the unfailing completion \cite{Bachmair89completionwithout}, superposition is a powerful calculus with 
a native support for equational reasoning. The calculus is parametrized by a simplification ordering on terms % and literals,
and uses it to constrain the applicability of inferences, with a significant impact on performance.

@@

In this work, we propose a machine learning system that learns to predict for an ATP
whether one precedence is going to lead to a faster proof search on a given problem than another.
Given a previously unseen problem, it can then be asked to recommend the best possible precedence for an ATP to run with.
Relying only on the logical structure of the problems, % for the learning, 
the system generalizes the knowledge about favourable precedences across problems with different signatures.

Our recommender uses a graph convolutional neural network \cite{kipf2017semisupervised}
to represent the problem structure. It learns from the ATP performance on selected problems
and \emph{pairs of randomly sampled} precedences. This information is then used to train
a \emph{symbol cost model}, which realizes the recommendation by simply sorting 
the problem's symbols according the obtained costs.

@@

The rest of this paper is organized as follows. \ldots
% explicit contributions (mixed with ``as follows'')?

\todo{Asi bych znovu (po abstraktu) nerikal, kolik nam vyjde. Spis jen ze sekce ta a ta je to tom, ze se tam evaluu s vampirem na TPTP. }
When coupled with the ATP Vampire and trained on problems from the TPTP library,
the recommender improves on the state-of-the art general-purpose precedence generating scheme
\texttt{invfreq} by more than \SI{4}{\percent}.

>>>>

Modern saturation-based Automatic Theorem Provers (ATPs) such as E \cite{SCV:CADE-2019}, SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09} 
or \Vampire{} \cite{DBLP:conf/cav/KovacsV13}
employ the superposition calculus \cite{DBLP:journals/logcom/BachmairG94,DBLP:books/el/RV01/NieuwenhuisR01} as their underlying inference system.
Integrating flavors of resolution \cite{DBLP:books/el/RV01/BachmairG01}, paramodulation \cite{Robinson1983}, and 
the unfailing completion \cite{Bachmair89completionwithout}, superposition is a powerful calculus with 
a native support for equational reasoning. The calculus is parameterized by a simplification ordering on terms % and literals,
and uses it to constrain the applicability of inferences, with a significant impact on performance.

@@

In this work, we propose a machine learning system that learns to predict for an ATP
whether one precedence is going to lead to a faster proof search on a given problem than another.
Given a previously unseen problem, it can then be asked to recommend the best possible precedence for an ATP to run with.
Relying only on the logical structure of the problems, % for the learning, 
the system generalizes the knowledge about favorable precedences across problems with different signatures.

Our recommender uses a graph convolutional neural network \cite{kipf2017semisupervised}
to represent the problem structure. It learns from the ATP performance on selected problems
and pairs of randomly sampled precedences. This information is then used to train
a \emph{symbol cost model}, which realizes the recommendation by simply sorting 
the problem's symbols according the obtained costs.

@@

The rest of this paper is organized as follows.
\Cref{sec:preliminaries} exposes the basic terminology used throughout the remaining sections.
\Cref{sec:architecture} proposes a structure of the precedence recommender that can be trained on pairs of symbol precedences.
\Cref{sec:evaluation} summarizes and discusses experiments performed
using an implementation of the precedence recommender.
\Cref{sec:related} compares the system proposed in this work with notable related works.
\Cref{sec:conclusion} concludes the investigation and outlines possible directions for future research.

### 2021-02-23 main.tex
Finalize the text for first submission

\section{Preliminaries}
\input{preliminaries}

@@

\section{Related work}

MS: my sami (PAAR paper) jsem si related workem. Budem se vuci tomu muset vymezit.
Jinak me ale nic moc bezprostredniho nenepada a nevim tedy, jestli se tedy primo
nutit do sekce s timto nazvem!

@@

\section{Conclusion}

@@

Generalization from trail acc to val acc may be improved.

Highest val accuracy does not transfer to highest ATP success rate. Generalization to ATP is sub-par.

Larger graphs

Joint model (trained jointly on predicate and function precedences)

Differentiate questions due to satiterations from questions due to success and failure

MS: Is there a knowledge transfer possible between related setups when we in vampire change
options (seemingly?) orthogonal to precedence (saturation algorithm, additional inference rules / reductions, AVATAR on/off,
clause selection strategies, etc.)?

KBO is a also parametrized by weights (along with the precedence). Could we learn those as well? (Jointly?)

Reinforcement learning will allow using a trained recommender to generate data from harder problems.

In our experiments, too much attention is spent on uninteresting examples.
We learn to order pairs of non-optimal precedences.
However, if the recommender ordered a non-optimal pair wrong, it could still produce the optimum precedence correctly.
Besides, we train from pairs of successful runs that only differ in number of iterations.

\todo[inline]{Idea for next project: Optimize precedence directly in a RL fashion: Generate precedence and evaluate with Vampire. If it is the best precedence seen so far, loss is 0 and gradient is 0. Otherwise ground truth is the best precedence seen so far on this problem and we use "differentiable ranking" to backpropagate. Possibly with some exploration, e.g. random swaps of pairs of symbols.}

>>>>

\section{Preliminaries}
\label{sec:preliminaries}
\input{preliminaries}

@@

\section{Related work}
\label{sec:related}

Our previous text \cite{DBLP:conf/cade/Bartek020} marked the initial investigation of applying techniques of \gls{ml}
to generating good symbol precedences.
The neural recommender uses a \gls{gnn} to model symbol costs,
while \cite{DBLP:conf/cade/Bartek020} used a linear combination of symbol features readily available in the \gls{atp} \Vampire{}.
The \gls{gnn}-based approach yields more performant precedences at the cost of longer training and preprocessing time.

\cite{Olsak2019} and \cite{Rawson2020} propose similar \gls{gnn} architectures to solve tasks on \gls{fol} problems.
They use the \glspl{gnn} to solve classification tasks such as premise selection.
While our system is trained on a proxy classification task,
the main task it is evaluated against is generating useful precedences.

@@

\section{Conclusion}
\label{sec:conclusion}

We have described a system that extracts useful symbol precedences from the graph representations of \gls{cnf} problems.
Comparison against a conventional symbol precedence heuristic shows that using a \gls{gcn}
to consider the whole structure of the input problem is beneficial.
\todo{FB: Should we boast that the reduction of ranking to training of differentiable element scores is innovative?}

@@

An analysis of the trained recommender could reveal new insights into how the symbol precedence influences the proof search,
and how a symbol precedence should be chosen in the context of a problem so that it is likely to lead to an efficient proof search.
This might lead to a design of a new precedence heuristic
that focuses relies on aspects of input problems
that have been overlooked so far in the context of generating symbol precedences.

While using the \gls{kbo} as our simplification ordering scheme of choice,
we restricted ourselves to generating the symbol precedences,
leaving the symbol weights fixed to the constant 1 (the default behavior of \Vampire{}).
Learning to recommend the symbol weights in addition to the symbol precedences would be interesting.
\todo{FB: Be more fancy in wording.}

Training both the predicate and the function precedence using a single \gls{gcn} would yield a recommender
that only needs one pass through a \gls{gcn} to generate both predicate and function precedences for a problem.
Moreover, the model would have the possibility to exploit the properties of the function precedence
when generating the predicate precedence, and vice versa.
Finally, higher training data efficiency could be achieved by considering all pairs of measured executions on a problem
in one training batch.
\todo{FB: Confusing?}

A more detailed analysis of the training process might yield insights into how the performance generalizes
from the training set to the validation set
and from proxy metrics, such as the loss and the accuracy, to the final solver performance.
Ultimately, this might reveal modifications that make the training more efficient and the final recommender more performant.

The training examples take the form of pairs of precedences,
where one of the precedences is preferable over the other.
In this work, the amount by which the quality of the precedences differs is ignored.
Taking the exact or estimated numbers of saturation loop iterations into account
might help the recommender to prioritize the more significant gains.

Applying techniques of reinforcement learning to focus the training data on the cases that yield high performance gains
could further improve the recommender performance.

We restricted our experiments to problems whose graph representation does not exceed \num{100000} nodes.
Evaluating the recommender on larger problems would be an interesting test of generalization.
Training on larger problems may be facilitated by employing a lossy graphification.

We performed all experiments with a fixed \gls{atp} \Vampire{} and a fixed configuration
we consider to be a reasonable baseline.
Introducing other prover configurations (namely other \glspl{sot})
or other provers could make the resulting recommender more robust
and possibly help extract more general knowledge about the structure of good symbol precedences.

Finally, the approach outlined in \cref{sec:architecture}
could be adjusted to solve tasks on \gls{fol} formulas other than recommending symbol precedences,
such as premise selection of given clause selection.
On the other hand, we could try to solve other tasks that involve generating permutations of arbitrary length.
\todo{FB: TSP? Probably not because it is too combinatorial.}

%KBO is a also parametrized by weights (along with the precedence). Could we learn those as well? (Jointly?)
